---
title: "WEEK-5 / Tidying and Transforming Data"
author: "Robert Green"
date: "2024-09-29"
output: html_document
---


```{r message=FALSE, warning=FALSE}
library(dplyr)
library(tidyr)
```


This data set was downloaded from the Bureau of Transportation Statistics website. 
A number of variables were filtered to reduce the size of the file. The file was
then downloaded to GITHUB. Later the file was read and assigned to 'airline_data'.

```{r}

airline_data <- read.csv("https://raw.githubusercontent.com/RGreen-sps/scaling-umbrella/refs/heads/main/T_ONTIME_REPORTING.csv")

str(airline_data)

```


The data set was filter further for 'UA', United Air Lines Inc., and 'WN', Southwest Airlines Co.
Additional filters includes the destination city names, and the origin city Chicago, IL.
New columns were created, 'on_time', 'delayed', and the carrier column renamed to 'airline.'
```{r include=FALSE}

airline_col_filter <- airline_data %>%
  filter(OP_UNIQUE_CARRIER %in% c("UA", "WN"),
         DEST_CITY_NAME %in% c("Denver, CO",
                               "New York, NY",
                               "Portland, OR",
                               "Sacramento, CA",
                               "St. Louis, MO"),
         ORIGIN_CITY_NAME == "Chicago, IL", 
         !is.na(ARR_DELAY)) %>% 
  select(OP_UNIQUE_CARRIER, DEST_CITY_NAME, ARR_DELAY) %>%
  group_by(OP_UNIQUE_CARRIER, DEST_CITY_NAME) %>% 
  summarise(on_time = sum(ARR_DELAY <= 0), delayed = sum(ARR_DELAY > 0)) %>% 
  arrange(DEST_CITY_NAME) %>% 
  rename(airline = OP_UNIQUE_CARRIER)


airline_col_filter

```
Pivot longer is used to combine the 'on_time' and 'delayed', into one column,
the status column.

```{r}

airline_add_status <- airline_col_filter |>
  pivot_longer(
    cols = on_time:delayed,
    names_to = "status",
    values_to = "NAME"
    )

airline_add_status

```
pivot wider is used here to turn the values in the DEST_CITY_NAME column to now 
act as column names for the new format.

```{r}

airline_compare <- airline_add_status |>
  pivot_wider(
    names_from = DEST_CITY_NAME,
    values_from = NAME
    )

airline_compare

```

Finally the data frame is written to a CSV file.

```{r}

write.csv(airline_compare, "airline_arr_delay_comp", row.names = FALSE)

```
